Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Paris version \YellowPaperVersionNumber}}

### 2023-12-07 Paper.tex
incorporate Paris EIPs

\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger\\ \smaller \textbf{{London version}}]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{London version \YellowPaperVersionNumber}}}
\author{
    Dr. Gavin Wood\\
    Founder, Ethereum \& Parity\\
    gavin@parity.io
}
\begin{document}

@@

Transactions are collated into blocks; blocks are chained together using a cryptographic hash as a means of reference. Blocks function as a journal, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big). They also punctuate the transaction series with incentives for nodes to \textit{mine}. This incentivisation takes place as a state-transition function, adding value to a nominated account.

Mining is the process of dedicating effort (working) to bolster one series of transactions (a block) over any other potential competitor block. It is achieved thanks to a cryptographically secure proof. This scheme is known as a proof-of-work and is discussed in detail in section \ref{ch:pow}.

Formally, we expand to:
\begin{eqnarray}
\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
B & \equiv & (..., (T_0, T_1, ...), ...) \\
\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Omega}{\Omega}(B, \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...)
\end{eqnarray}

Where \hyperlink{Omega}{$\Omega$} is the block-finalisation state transition function (a function that rewards a nominated party); \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components; and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.

@@

The scheme we use in order to generate consensus is a simplified version of the GHOST protocol introduced by \cite{cryptoeprint:2013:881}. This process is described in detail in section \ref{ch:ghost}.

Sometimes, a path follows a new protocol from a particular height (block number).
This document describes the \textit{London}, \textit{Arrow Glacier}, and \textit{Gray Glacier} versions of the protocol.\footnote{Note that the \textit{Arrow Glacier} and \textit{Gray Glacier} forks only affected the \hyperlink{difficulty_bomb}{"difficulty bomb"} described in section \ref{subsec:The_Block}.}
In order to follow back the history of a path, one must reference multiple versions of this document.
Here are the block numbers of protocol updates on the Ethereum main network:
\par
\begin{center}
\begin{tabular}{lr}
\toprule
Name & First Block Number \\
\midrule
$F_{\mathrm{Homestead}}$         &  1150000 \\
$F_{\mathrm{Tangerine Whistle}}$ &  2463000 \\
$F_{\mathrm{Spurious Dragon}}$   &  2675000 \\
$F_{\mathrm{Byzantium}}$         &  4370000 \\
$F_{\mathrm{Constantinople}}$    &  7280000 \\
$F_{\mathrm{Petersburg}}$        &  7280000 \\
$F_{\mathrm{Istanbul}}$          &  9069000 \\
$F_{\mathrm{Muir Glacier}}$      &  9200000 \\
$F_{\mathrm{Berlin}}$            & 12244000 \\
$F_{\mathrm{London}}$            & 12965000 \\
$F_{\mathrm{Arrow Glacier}}$     & 13773000 \\
$F_{\mathrm{Gray Glacier}}$      & 15050000 \\
\bottomrule
\end{tabular}
\end{center}
\par

@@

The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language_in_English\#Aunt/Uncle}}). The block header contains several pieces of information:

@@

\begin{description}
\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
\item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally \hyperlink{block_number_H__i}{$H_{\mathrm{i}}$}.
\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
\item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.
\item[baseFeePerGas]\linkdest{block_baseFeePerGas_H__f}{} A scalar value equal to the amount of wei that is burned for each unit of gas consumed; formally \hyperlink{H__f}{$H_{\mathrm{f}}$}.
\end{description}
\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$. Formally, we can refer to a block $B$:
\begin{equation}
B \equiv (B_{\mathrm{H}}, B_{\mathbf{T}}, B_{\mathbf{U}})
\end{equation}

@@

\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
\begin{equation}
\begin{array}[t]{lclc}
\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{KEC}(\mathtt{RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p_{\mathrm{T}}(i, B_{\mathbf{T}}[i])\}) & \wedge \\
\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p_{\mathrm{R}}(i, B_{\mathbf{R}}[i])\}) & \wedge \\
\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
\end{array}
\end{equation}
where $p_{\mathrm{T}}(k, v)$ and $p_{\mathrm{R}}(k, v)$ are pairwise RLP transformations, but with a special treatment for EIP-2718 transactions:
\begin{equation}
p_{\mathrm{T}}(k, T) \equiv \left( \mathtt{RLP}(k), \begin{cases}
\mathtt{RLP}(\hyperlink{L_transaction}{L_{\mathrm{T}}}(T)) & \text{if} \quad T_{\mathrm{x}} = 0 \\
(T_{\mathrm{x}}) \cdot \mathtt{RLP}(L_{\mathrm{T}}(T)) & \text{otherwise}
\end{cases}
\right)
\end{equation}
and
\begin{equation}
p_{\mathrm{R}}(k, R) \equiv \left( \mathtt{RLP}(k), \begin{cases}
\mathtt{RLP}(\hyperlink{L__R}{L_{\mathrm{R}}}(R)) & \text{if} \quad R_{\mathrm{x}} = 0 \\
(R_{\mathrm{x}}) \cdot \mathtt{RLP}(L_{\mathrm{R}}(R)) & \text{otherwise}
\end{cases}
\right)
\end{equation}
($\cdot$ is the concatenation of byte arrays).

@@

\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\linkdest{L__B}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}\linkdest{L__B}The function $L_{\mathrm{B}}$ and $L_{\mathrm{H}}$ are the preparation functions for a block and block header respectively.
We assert the types and order of the structure for when the RLP transformation is required:
\begin{eqnarray}
\quad L_{\mathrm{H}}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}}, H_{\mathrm{f}} \; )\end{array} \\
\quad L_{\mathrm{B}}(B) & \equiv & \big( L_{\mathrm{H}}(B_{\mathrm{H}}), \widetilde{L}_{\mathrm{T}}^*(B_{\mathbf{T}}), L_{\mathrm{H}}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
\end{eqnarray}
where $\widetilde{L}_{\mathrm{T}}$ takes a special care of EIP-2718 transactions:
\begin{equation}
\widetilde{L}_{\mathrm{T}}(T) = \begin{cases}
\hyperlink{L_transaction}{L_{\mathrm{T}}}(T) & \text{if} \quad T_{\mathrm{x}} = 0 \\
(T_{\mathrm{x}}) \cdot \mathtt{RLP}(L_{\mathrm{T}}(T)) & \text{otherwise}
\end{cases}
\end{equation}
\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}with $\widetilde{L}_{\mathrm{T}}^*$ and $L_{\mathrm{H}}^*$ being element-wise sequence transformations, thus:
\begin{equation}
f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
\end{equation}

The component types are defined thus:
\begin{equation}
\begin{array}[t]{lclclcl}
\hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{o}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
\hyperlink{new_state_H__r}{H_{\mathrm{r}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & \hyperlink{Receipts_Root_H__e}{H_{\mathrm{e}}} \in \mathbb{B}_{32} & \wedge \\
\hyperlink{logs_Bloom_filter_H__b}{H_{\mathrm{b}}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{N} & \wedge & \hyperlink{block_number_H__i}{H_{\mathrm{i}}} \in \mathbb{N} & \wedge \\
\hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{N} & \wedge & H_{\mathrm{g}} \in \mathbb{N} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{N}_{256} & \wedge \\
\hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{m}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8} & \wedge \\
\hyperlink{block_baseFeePerGas_H__f}{H_{\mathrm{f}}} \in \mathbb{N}

@@

\hypertarget{block_difficulty_H__d}{}\linkdest{H__d}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
\begin{equation}
D(H) \equiv \begin{dcases}
2^{34} & \text{if} \quad H_{\mathrm{i}} = 0\\
\text{max}\!\left(\mindifficulty, {P(H)_{\mathrm{H}}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
\end{dcases}
\end{equation}
where:
\begin{align}
\mindifficulty &\equiv 2^{17} \\
\diffadjustment &\equiv \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{d}}}{2048}\right\rfloor \\
\homesteadmod &\equiv \text{max}\left( y - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{\mathrm{H}}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right) \\
y &\equiv \begin{cases}
1 & \text{if} \quad \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
2 & \text{otherwise}
\end{cases} \\
\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - \kappa, 0)
\end{align}
\begin{equation}
\kappa \equiv \begin{cases} 
  3000000  & \text{if} \quad F_{\mathrm{Byzantium}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Constantinople}} \\
  5000000  & \text{if} \quad F_{\mathrm{Constantinople}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Muir Glacier}} \\
  9000000  & \text{if} \quad F_{\mathrm{Muir Glacier}} \leqslant H_{\mathrm{i}} < F_{\mathrm{London}} \\
  9700000  & \text{if} \quad F_{\mathrm{London}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Arrow Glacier}} \\
 10700000  & \text{if} \quad F_{\mathrm{Arrow Glacier}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Gray Glacier}} \\
 11400000  & \text{if} \quad H_{\mathrm{i}} \geqslant F_{\mathrm{Gray Glacier}} \\
\end{cases}
\end{equation}

\hypertarget{difficulty_bomb}{}The \textit{Homestead} difficulty parameter, $\homesteadmod$, is used to affect a dynamic homeostasis of time between blocks, as the time between blocks varies, as discussed below, as implemented in EIP-2 by \cite{EIP-2}.
In the Homestead release, the exponential difficulty symbol, $\expdiffsymb$ causes the difficulty to slowly increase (every 100,000 blocks) at an exponential rate, and thus increasing the block time difference, and putting time pressure on transitioning to proof-of-stake.
This effect, known as the "difficulty bomb", or "ice age", was explained in EIP-649 by \cite{EIP-649} and delayed and implemented earlier in EIP-2.
$\homesteadmod$ was also modified in EIP-100 with the use of $x$, the adjustment factor above, and the denominator 9, in order to target the mean block time including uncle blocks by \cite{EIP-100}.
In the \textit{Byzantium} release, with EIP-649, the ice age was delayed by creating a fake block number, $H'_{\mathrm{i}}$, which is obtained by subtracting three million from the actual block number,
which in other words reduced $\expdiffsymb$ and the time difference between blocks, in order to allow more time to develop proof-of-stake and preventing the network from "freezing" up.
Subsequently, EIP-1234 by \cite{EIP-1234}, EIP-2384 by \cite{EIP-2384}, EIP-3554 by \cite{EIP-3554}, EIP-4345 by \cite{EIP-4345}, and EIP-5133 by \cite{EIP-5133} increased the subtrahend $\kappa$.

@@

This mechanism enforces a homeostasis in terms of the time between blocks; a smaller period between the last two blocks results in an increase in the difficulty level and thus additional computation required, lengthening the likely next period. Conversely, if the period is too large, the difficulty, and expected time to the next block, is reduced.

The nonce, \linkdest{H__n}\hyperlink{block_nonce_H__n}{$H_{\mathrm{n}}$}, must satisfy the relations:
\begin{equation}
n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \quad \wedge \quad m = H_{\mathrm{m}}
\end{equation}
with $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$.

\hypertarget{block_header_without_nonce_and_mixmash_h__cancel_n}{}\linkdest{H_cancel_n}Where $H_{\hcancel{n}}$ is the new block's header $H$, but \textit{without} the nonce and mix-hash components, $\mathbf{d}$ being the current DAG, a large data set needed to compute the mix-hash, and $\mathtt{PoW}$ is the proof-of-work function (see section \ref{ch:pow}): this evaluates to an array with the first item being the mix-hash, to prove that a correct DAG has been used, and the second item being a pseudo-random number cryptographically dependent on $H$ and $\mathbf{d}$. Given an approximately uniform distribution in the range $[0, 2^{64})$, the expected time to find a solution is proportional to the difficulty, $H_{\mathrm{d}}$.

>>>>

\title[Ethereum: A Secure Decentralised Generalised Transaction Ledger\\ \smaller \textbf{{Paris version}}]{Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Paris version \YellowPaperVersionNumber}}}
\author{
    Dr. Gavin Wood\\
    Founder, Ethereum \& Parity\\
    gavin@parity.io
}
\begin{document}

@@

Transactions are collated into blocks; blocks are chained together using a cryptographic hash as a means of reference. Blocks function as a journal, recording a series of transactions together with the previous block and an identifier for the final state (though do not store the final state itself---that would be far too big).

Formally, we expand to:
\begin{eqnarray}
\boldsymbol{\sigma}_{t+1} & \equiv & \hyperlink{Pi}{\Pi}(\boldsymbol{\sigma}_{t}, B) \\
B & \equiv & (..., (T_0, T_1, ...), ...) \\
\Pi(\boldsymbol{\sigma}, B) & \equiv & \hyperlink{Upsilon}{\Upsilon}(\Upsilon(\boldsymbol{\sigma}, T_0), T_1) ...
\end{eqnarray}

Where \hyperlink{block}{$B$} is this block, which includes a series of transactions amongst some other components and $\hyperlink{Pi}{\Pi}$ is the block-level state-transition function.

@@

Since the \textit{Paris} hard fork, reaching consensus on new blocks is managed by a protocol called the \textit{Beacon Chain}.
It is known as the \textit{consensus layer} of Ethereum, and it defines the rules for determining the canonical history of Ethereum blocks.
This document describes the \textit{execution layer} of Ethereum.
The execution layer defines the rules for interacting with and updating the state of the Ethereum virtual machine.
The consensus layer is described in greater detail in the \href{https://github.com/ethereum/consensus-specs}{consensus specifications}.
How the consensus layer is used to determine the canonical state of Ethereum is discussed in section \ref{ch:blocktree_to_blockchain}.

Sometimes, a path follows a new protocol from a particular height (block number).
This document describes the \textit{Paris} version.

In order to follow back the history of a path, one must reference multiple versions of this document.
Here are the block numbers of protocol updates on the Ethereum main network:\footnote{Note that while the Paris fork activated at block 15,537,393, the trigger was not the block number, but rather reaching a specified \textit{total difficulty}. The trigger for the \textit{Paris} hard fork is discussed in greater detail in section \ref{ch:pos_transition}.}
\par
\begin{center}
\begin{tabular}{lr}
\toprule
Name & First Block Number \\
\midrule
$F_{\mathrm{Homestead}}$         &  1150000 \\
$F_{\mathrm{Tangerine Whistle}}$ &  2463000 \\
$F_{\mathrm{Spurious Dragon}}$   &  2675000 \\
$F_{\mathrm{Byzantium}}$         &  4370000 \\
$F_{\mathrm{Constantinople}}$    &  7280000 \\
$F_{\mathrm{Petersburg}}$        &  7280000 \\
$F_{\mathrm{Istanbul}}$          &  9069000 \\
$F_{\mathrm{Muir Glacier}}$      &  9200000 \\
$F_{\mathrm{Berlin}}$            & 12244000 \\
$F_{\mathrm{London}}$            & 12965000 \\
$F_{\mathrm{Arrow Glacier}}$     & 13773000 \\
$F_{\mathrm{Gray Glacier}}$      & 15050000 \\
$F_{\mathrm{Paris}}$      & 15537393 \\
\bottomrule
\end{tabular}
\end{center}
\par

@@

The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with information corresponding to the comprised transactions, $\mathbf{T}$,\hypertarget{ommerheaders}{} and a now deprecated property $\mathbf{U}$ that prior to the \textit{Paris} hard fork contained headers equal to the present block's parent's parent (such blocks are known as \textit{ommers}\footnote{\textit{ommer} is a gender-neutral term to mean ``sibling of parent''; see \url{https://nonbinary.miraheze.org/wiki/Gender_neutral_language_in_English\#Aunt/Uncle}}). The block header contains several pieces of information:

@@

\begin{description}
\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
\item[ommersHash] A 256-bit hash field that is now deprecated due the replacement of proof of work consensus. It is now to a constant, \texttt{Keccak256(RLP([]))}; formally $H_{\mathrm{o}}$.
\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which priority fees from this block are transferred; formally $H_{\mathrm{c}}$.
\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
\item[difficulty] A scalar field that is now deprecated due to the replacement of proof of work consensus. It is set to 0; formally $H_{\mathrm{d}}$.
\item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally \hyperlink{block_number_H__i}{$H_{\mathrm{i}}$}.
\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
\item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
\item[prevRandao]\linkdest{prevRandao_H__a}{}\linkdest{H__a} the latest RANDAO mix\footnote{RANDAO is a pseudorandom value generated by validators on the Ethereum consensus layer. Refer to the consensus layer specs (\url{https://github.com/ethereum/consensus-specs}) for more detail on RANDAO.} of the post beacon state of the previous block; formally $H_{\mathrm{a}}$.
\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value that is now deprecated due to the replacment of proof of work consensus. It is set to \texttt{0x0000000000000000}; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.
\item[baseFeePerGas]\linkdest{block_baseFeePerGas_H__f}{} A scalar value equal to the amount of wei that is burned for each unit of gas consumed; formally \hyperlink{H__f}{$H_{\mathrm{f}}$}.
\end{description}
\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are a series of transactions, $B_{\mathbf{T}}$, and an empty array which was previously reserved for ommer block headers, $B_{\mathbf{U}}$. Formally, we can refer to a block $B$:
\begin{equation}
B \equiv (B_{\mathrm{H}}, B_{\mathbf{T}}, B_{\mathbf{U}})
\end{equation}

@@

\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: its ommers field $B_{\mathbf{U}}$ must be an empty array and it must be internally consistent with the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
\begin{equation}
\begin{array}[t]{lclc}
B_{\mathbf{U}} &\equiv& [] & \wedge \\
\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p_{\mathrm{T}}(i, B_{\mathbf{T}}[i])\}) & \wedge \\
\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p_{\mathrm{R}}(i, B_{\mathbf{R}}[i])\}) & \wedge \\
\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
\end{array}
\end{equation}
where $p_{\mathrm{T}}(k, v)$ and $p_{\mathrm{R}}(k, v)$ are pairwise RLP transformations, but with a special treatment for EIP-2718 transactions:
\begin{equation}
p_{\mathrm{T}}(k, T) \equiv \left( \mathtt{RLP}(k), \begin{cases}
\mathtt{RLP}(\hyperlink{L_transaction}{L_{\mathrm{T}}}(T)) & \text{if} \quad T_{\mathrm{x}} = 0 \\
(T_{\mathrm{x}}) \cdot \mathtt{RLP}(L_{\mathrm{T}}(T)) & \text{otherwise}
\end{cases}
\right)
\end{equation}
and
\begin{equation}
p_{\mathrm{R}}(k, R) \equiv \left( \mathtt{RLP}(k), \begin{cases}
\mathtt{RLP}(\hyperlink{L__R}{L_{\mathrm{R}}}(R)) & \text{if} \quad R_{\mathrm{x}} = 0 \\
(R_{\mathrm{x}}) \cdot \mathtt{RLP}(L_{\mathrm{R}}(R)) & \text{otherwise}
\end{cases}
\right)
\end{equation}
($\cdot$ is the concatenation of byte arrays).

@@

\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\linkdest{L__B}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}\linkdest{L__B}The function $L_{\mathrm{B}}$ and $L_{\mathrm{H}}$ are the preparation functions for a block and block header respectively.
We assert the types and order of the structure for when the RLP transformation is required:
\begin{eqnarray}
\quad L_{\mathrm{H}}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{a}}, H_{\mathrm{n}}, H_{\mathrm{f}} \; )\end{array} \\
\quad L_{\mathrm{B}}(B) & \equiv & \big( L_{\mathrm{H}}(B_{\mathrm{H}}), \widetilde{L}_{\mathrm{T}}^*(B_{\mathbf{T}}), L_{\mathrm{H}}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
\end{eqnarray}
where $\widetilde{L}_{\mathrm{T}}$ takes a special care of EIP-2718 transactions:
\begin{equation}
\widetilde{L}_{\mathrm{T}}(T) = \begin{cases}
\hyperlink{L_transaction}{L_{\mathrm{T}}}(T) & \text{if} \quad T_{\mathrm{x}} = 0 \\
(T_{\mathrm{x}}) \cdot \mathtt{RLP}(L_{\mathrm{T}}(T)) & \text{otherwise}
\end{cases}
\end{equation}
\hypertarget{general_element_wise_sequence_transformation_f_pow_asterisk}{}with $\widetilde{L}_{\mathrm{T}}^*$ and $L_{\mathrm{H}}^*$ being element-wise sequence transformations, thus:
\begin{equation}
f^*\big( (x_0, x_1, ...) \big) \equiv \big( f(x_0), f(x_1), ... \big) \quad \text{for any function} \; f
\end{equation}

The component types are defined thus:
\begin{equation}
\begin{array}[t]{lclclcl}
\hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{o}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{c}} \in \mathbb{B}_{20} & \wedge \\
\hyperlink{new_state_H__r}{H_{\mathrm{r}}} \in \mathbb{B}_{32} & \wedge & H_{\mathrm{t}} \in \mathbb{B}_{32} & \wedge & \hyperlink{Receipts_Root_H__e}{H_{\mathrm{e}}} \in \mathbb{B}_{32} & \wedge \\
\hyperlink{logs_Bloom_filter_H__b}{H_{\mathrm{b}}} \in \mathbb{B}_{256} & \wedge & H_{\mathrm{d}} \in \mathbb{N} & \wedge & \hyperlink{block_number_H__i}{H_{\mathrm{i}}} \in \mathbb{N} & \wedge \\
\hyperlink{block_gas_limit_H__l}{H_{\mathrm{l}}} \in \mathbb{N} & \wedge & H_{\mathrm{g}} \in \mathbb{N} & \wedge & \hyperlink{block_timestamp_H__s}{H_{\mathrm{s}}} \in \mathbb{N}_{256} & \wedge \\
\hyperlink{block_extraData_H__x}{H_{\mathrm{x}}} \in \mathbb{B} & \wedge & H_{\mathrm{a}} \in \mathbb{B}_{32} & \wedge & \hyperlink{block_nonce_H__n}{H_{\mathrm{n}}} \in \mathbb{B}_{8} & \wedge \\
\hyperlink{block_baseFeePerGas_H__f}{H_{\mathrm{f}}} \in \mathbb{N}

@@

@@

The \textit{Paris} hard fork changed Ethereum's consensus from proof of work to proof of stake, and thus deprecated many block header properties related to proof of work. These deprecated properties include \textbf{nonce} ($H_{\mathbf{n}}$), \textbf{ommersHash} ($H_{\mathbf{o}}$), \textbf{difficulty} ($H_{\mathbf{d}}$), and \textbf{mixHash} ($H_{\mathbf{m}}$).

\textbf{mixHash} has been replaced with a new field \textbf{prevRandao} ($H_{\mathbf{a}}$). The other header fields related to proof of work have been replaced with constants:

\begin{eqnarray}
  H_{\mathrm{o}} & \equiv & \mathtt{KEC}( \mathtt{RLP}([])) \\
  H_{\mathrm{d}} & \equiv  & 0 \\
  H_{\mathrm{n}}  & \equiv & \texttt{0x0000000000000000}
\end{eqnarray}

The value of \textbf{prevRandao} must be determined using information from the Beacon Chain.
While the details of generating the \textit{RANDAO} value on the Beacon Chain is beyond the scope of this paper, we refer to the expected \textit{RANDAO} value for the previous block as $\mathtt{PREVRANDAO}()$.
